---
title: "Examine IRR"
author: "Rex Macey"
date: "January 24, 2016"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(stringr)
library(plyr)
library(lubridate)
library(randomForest)
library(reshape2)
library(formattable)
setwd(Sys.getenv("LENDINGCLUB_DIR"))  # this gets the LENDINGCLUB_DIR variable defined in the .rProfile file
df<-readRDS("LoanStatsAll.rds")
df2<-df # used for all loans
```
(Modified 1/24/2016 to add weighted means.  Weighting is by funding amount.)
The number of loans (rows of data) is `r accounting(nrow(df),0)`.

We need to consider completed loans. It's not fair to include a loan initiated 1 year ago that defaulted.
Doing so would include bad loans from recent periods but not loans that turn out to be good.
We filter for loans that are have a term of 3 years initiated at least 3 years ago or a term of 5 years
initiated at least 5 years ago.

```{r message=FALSE, warning=FALSE}
max_issue_d<-max(df$issue_d,na.rm=TRUE)
idx<- (df2$issue_d < (max_issue_d - 3) & df$term == " 36 months") | (df2$issue_d < (max_issue_d - 5) & df$term == " 60 months")
df<-df[idx,]
```

Now we are down to a much smaller number (`r accounting(nrow(df),0)`) of loans (about `r paste0(round(100*nrow(df)/nrow(df2),1),"%")` of the entire data set).  Does it seem correct that
LC has initiated 10x loans since these were complete.

The loans have terms of either 3 or 5 years.  Below we see that 3 Year loans dominate completed loans
```{r message=FALSE, warning=FALSE}
tbl<-rbind(summary(df$term),round(100*summary(df$term)/nrow(df),0))[,2:3]
rownames(tbl)<-c("Count","Pct")
tbl
```
It appears LC is making many more 5 year loans now based on the table below which considers all loans.
```{r message=FALSE, warning=FALSE}
tbl<-rbind(summary(df2$term),round(100*summary(df2$term)/nrow(df2),0))[,2:3]
rownames(tbl)<-c("Count","Pct")
tbl
```
So a model built on completed loans only has `r accounting(summary(df$term)[3],0)` five-year loans. One may want to limit the model to three year loans.

Now we turn to the status of the loans as classified by LC. First we list the status and then show how many "completed" loans fall in to each category.
```{r message=FALSE, warning=FALSE}
status_levels<-levels(df$loan_status)
status_levels
summary(df$loan_status)
```
It's odd that a a handful of the loans we thought would be completed are either late (4)  or in default (1).
To be careful we only use loans with the following codes
```{r message=FALSE, warning=FALSE}
complete_status<-status_levels[c(2,5,7,8)]
complete_status
idx<-df$loan_status %in% complete_status
df<-df[idx,]
loan.status.dist<-data.frame(table(df$loan_status)) #distribution of by loan status
loan.status.dist$PctTotal<-round(100*loan.status.dist$Freq/sum(loan.status.dist$Freq),1)
loan.status.dist<-loan.status.dist[loan.status.dist$Freq>0,]
loan.status.dist

# Calculate or lookup IRR
if (FALSE) {
    source("IRR.R")
    df$irr<-NA
    for(i in 1:nrow(df)){
        df$irr[i]<-IRR(df$total_pymnt[i],df$total_rec_late_fee[i],df$collection_recovery_fee[i],df$installment[i],df$funded_amnt[i],df$issue_d[i])
    }
    saveRDS(df,file=paste0(getwd(),"/","irr.rds"))    
} else {
    df<-readRDS(paste0(getwd(),"/","irr.rds"))
}
summary(df$irr) # negative skew
boxplot(df$irr,main="Boxplot of IRR",ylab="IRR",xlab="All Observations")

# IRR by loan status
irr<-df$irr
meanbystatus <- ddply(df, .(loan_status),summarise,
                      MeanIRR=round(100*mean(irr,na.rm=T),1),
                      WtMeanIRR=round(100*weighted.mean(irr,funded_amnt,na.rm=TRUE),1),
                      MedianIRR=round(100*median(irr,na.rm=T),1),
                      Count=length(irr),
                      FreqPct=round(100*length(irr)/length(df[,1]),1))
print(meanbystatus)

meanbygrade <- ddply(df, .(grade),summarise,
                      MeanIRR=round(100*mean(irr,na.rm=T),1),
                      WtMeanIRR=round(100*weighted.mean(irr,funded_amnt,na.rm=TRUE),1),
                      MedianIRR=round(100*median(irr,na.rm=T),1),
                      Count=length(irr),
                      FreqPct=round(100*length(irr)/length(df[,1]),1))
print(meanbygrade)

meanbyyear <- ddply(df, .(year_issued),summarise,
                      MeanIRR=round(100*mean(irr,na.rm=TRUE),1),
                      WtMeanIRR=round(100*weighted.mean(irr,funded_amnt,na.rm=TRUE),1),
                      MedianIRR=round(100*median(irr,na.rm=T),1),
                      Count=length(irr),
                      FreqPct=round(100*length(irr)/length(df[,1]),1))
print(meanbyyear)
```