---
title: "ExamineCF.rmd"
author: "Rex Macey"
date: "January 24, 2016"
output: html_document
---
```{r message=FALSE, warning=FALSE}
library(stringr)
library(plyr)
library(lubridate)
library(randomForest)
library(reshape2)
library(formattable)
setwd(Sys.getenv("LENDINGCLUB_DIR"))  # this gets the LENDINGCLUB_DIR variable defined in the .rProfile file
df<-readRDS("LoanStatsAll.rds")
df2<-df # used for all loans
max_issue_d<-max(df$issue_d,na.rm=TRUE)
idx<- (df2$issue_d < (max_issue_d - 3) & df$term == " 36 months") | (df2$issue_d < (max_issue_d - 5) & df$term == " 60 months")
df<-df[idx,]
```
Let's look at cash flows to make sure we understand them.  Here are some of the relevant fields.

```{r message=FALSE, warning=FALSE}
amnt_cols<-strsplit("id,funded_amnt,funded_amnt_inv,total_pymnt,total_pymnt_inv,out_prncp,out_prncp_inv,total_rec_late_fee,installment,collection_recovery_fee,recoveries,total_rec_int,total_rec_prncp,loan_status,term,int_rate",split=",")[[1]]
df.cf<-df[,amnt_cols]
head(df.cf[idx,])
```
How does funded_amnt differ from funded_amnt_inv?
How many are different (#/%)?
Let's look at a few of the loans with differences, especially the larger differences.
I think the difference is the amount of the loan funded by LC itself.  
```{r message=FALSE, warning=FALSE}
summary(df.cf$funded_amnt-df.cf$funded_amnt_inv) # always non-negative
idx<-df.cf$funded_amnt>df.cf$funded_amnt_inv
print(paste0(sum(idx)," / ",round(100*sum(idx)/nrow(df.cf),0),"%"))
head(df.cf[idx,])
df.cf.sort<-df.cf[order(-df.cf$funded_amnt+df.cf$funded_amnt_inv),]
head(df.cf.sort,20)
rm(df.cf.sort)
```

Let's look at total_pymnt v total_pymnt_inv.
Note that total_pymnt exceeds total_pymnt_inv more often than funded_amnt exceeds funded_amnt_inv.
Perhaps this is because LC may receive some payments that do not go to investors?
```{r message=FALSE, warning=FALSE}
summary(df.cf$total_pymnt-df.cf$total_pymnt_inv) # always non-negative
idx<-df.cf$total_pymnt>df.cf$total_pymnt_inv
print(paste0(sum(idx)," / ",round(100*sum(idx)/nrow(df.cf),0),"%"))
```

What consistutes total_pymnt?
We can see it is total_rec_int+total_rec_prncp+total_rec_late_fee+recoveries
```{r message=FALSE, warning=FALSE}
x<-df.cf$total_rec_int+df.cf$total_rec_prncp+df.cf$total_rec_late_fee+df.cf$recoveries
idx<-abs(df.cf$total_pymnt - x) > 0.01
print(paste0(sum(idx)," / ",round(100*sum(idx)/nrow(df.cf),0),"%"))
```

What consistutes total_pymnt_inv?
Q) Is it a simple proportion related to funded_amnt_inv/funded_amnt
A) No, though it does a good job for more than 50% of the loans.  Errors on both sides of zero
```{r message=FALSE, warning=FALSE}
test<-df.cf$total_pymnt_inv - df.cf$total_pymnt * df.cf$funded_amnt_inv / df.cf$funded_amnt
summary(test)
df.cf.sort<-df.cf
df.cf.sort$test<-test
df.cf.sort<-df.cf.sort[order(-df.cf.sort$test),]
head(df.cf.sort)
tail(df.cf.sort)
```

Q)Are the loans with unexplained total_pymnt_inv related to loan status?
First a summary of the status for all loans, then for those with errors
A) All are related to charge offs, but not all charge offs are unexplained
```{r message=FALSE, warning=FALSE}
summary(df.cf.sort$loan_status)
idx<-abs(df.cf.sort$test>=1) # not worried about issues less than $1
df.unexp<-df.cf.sort[idx,]
summary(df.unexp$loan_status)
```

Is the unexplained amount of total_pymnt_inv related to the recovery?
Not according to the graph
```{r message=FALSE, warning=FALSE}
idx<-df.cf.sort$test>=1
plot(df.cf.sort$recoveries[idx],df.cf.sort$test[idx],
     main="Trying to explain total_pymnt_inv with recoveries",xlab="Recoveries",ylab="Unexplained")
```

If we can't exactly explain total_pymnt_inv, is that a problem?  Let's start with the hypothesis that we investors receive total_pymnt_inv and that someone else receives the rest of total_pymnt.  As long as the investors get their fair share, we can focus on total_pymnt and not total_pymnt_inv.

Well sometimes they don't.

```{r message=FALSE, warning=FALSE}
inv_pct<-df.cf$funded_amnt_inv/df.cf$funded_amnt # investors % of the loan
inv_excess<-df.cf$total_pymnt_inv - df.cf$total_pymnt*inv_pct # +/- dollars 
summary(inv_excess)
```

Finally, let's just look at total funded v total payments. 
```{r message=FALSE, warning=FALSE}
paymentsbystatus <- ddply(df, .(loan_status),summarise,
                      Funded=comma(sum(funded_amnt,na.rm=T),0),
                      Paid=comma(sum(total_pymnt,na.rm=TRUE),0),
                      FundedInv=comma(sum(funded_amnt_inv,na.rm=T),0),
                      PaidInv=comma(sum(total_pymnt_inv,na.rm=TRUE),0),
                      Count=comma(length(funded_amnt),0),
                      FreqPct=percent(length(funded_amnt)/length(df[,1]),1))
print(paymentsbystatus)

paymentsbyyear <- ddply(df, .(year_issued),summarise,
                      Funded=comma(sum(funded_amnt,na.rm=T),0),
                      Paid=comma(sum(total_pymnt,na.rm=TRUE),0),
                      FundedInv=comma(sum(funded_amnt_inv,na.rm=T),0),
                      PaidInv=comma(sum(total_pymnt_inv,na.rm=TRUE),0),
                      Count=comma(length(funded_amnt),0),
                      FreqPct=percent(length(funded_amnt)/length(df[,1]),1))
print(paymentsbyyear)

paymentsbygrade <- ddply(df, .(grade),summarise,
                      Funded=comma(sum(funded_amnt,na.rm=T),0),
                      Paid=comma(sum(total_pymnt,na.rm=TRUE),0),
                      FundedInv=comma(sum(funded_amnt_inv,na.rm=T),0),
                      PaidInv=comma(sum(total_pymnt_inv,na.rm=TRUE),0),
                      Count=comma(length(funded_amnt),0),
                      FreqPct=percent(length(funded_amnt)/length(df[,1]),1))
print(paymentsbygrade)
```

